---
title: "Gov 50 Final Project"
author: "Nina Howe-Goldstein"
description: "My final project: D.C. copterspotting"
output:
  distill::distill_article:
    self_contained: false
---

```{r}
library(tidyverse)
library(lubridate)
copterspotter <- read.csv("copterspotter 2 - tidyer - copterspotter.csv")
copterspotter$Date <- ymd(copterspotter$Date)
copterspotter_clean <- copterspotter |>
  filter(Lon <= -70) |>
  filter(Lat <= 39)
```

Introduction:

Helicopters are everywhere in my hometown of Washington, D.C. On any given day, four different police departments, news choppers, medivacs, the Park Police, the Army, the Navy, the Air Force, the National Guard, Homeland Security, and the elite Marine squadron that transports the president might be flying around the District. Subject to little scrutiny because of their broad national security mandate – most don’t even have to turn on their transponders – noisy helicopters become a nuisance and a spectacle for bored Washingtonians. During the pandemic, one irritated teleworker founded Helicopters of D.C., a helicopter tracking site that creates a daily user-submitted map of sightings across the DMV. 

I love Helicopters of D.C. I’ve registered more than 300 helicopter sightings with them – including one incredible afternoon at the airport when my flight was delayed by six hours, and I parked myself by the window to watch takeoffs and landings at Joint Base Anacostia-Bolling across the river. (That pales in comparison to the site’s highest-ranked spotters, including an engineer at the Department of Transportation with a great office view and more than 5000 spots to her name.)

But the original purpose of Helicopters of D.C. was, in part, to track noise pollution. When helicopters fly low – which they almost always do, without consequence – they are extremely loud and obnoxious. To address locals’ complaints, D.C.’s non-voting Congressional delegate Eleanor Holmes Norton (who is perhaps most famous for putting out a bizarre statement denying involvement in the escape of six zebras from a Maryland zoo in 2021) introduced a series of bills in the summer of 2022 to standardize helicopter regulations. Norton’s bills would require helicopters to fly at the maximum altitude permitted by the Federal Aviation Administration, with limited exceptions, and instituted a trial noise complaint system for D.C. residents. Although Norton’s bills are still working their way through Congress, given that the lawmakers responsible for voting on the new regulations are also affected by D.C. copter noise pollution, I wondered if there might have been a reduction in obnoxious helicopter use after each bill’s introduction.

My hypothesis was that Helicopters of D.C.’s sighting data from 2022 would show a slight reduction in spottings, indicative of a reduction in obnoxious helicopter use — since a user is more likely to be prompted by the noise of a low-flying helicopter to look up and take a picture. (Based on anecdotal experience from living in D.C., I also guessed that when broken down by operator, the data would show no reduction in obnoxious use by the Air Force 1st Helicopter Squadron. Their UH-1N Hueys are some of the worst offenders.) 


Data:

In its end-of-year report for 2022, Helicopters of D.C. also published an anonymized table of the full year’s data. I quickly learned that it was a huge pain to wrangle: someone had submitted a few helicopter spots from Madrid as a joke, which threw off my first data visualization until I could `select` them out, and I had to install the Lubridate package to change the `Date` column from character data to dates that R would recognize. 

To make matters worse, I had hoped to do a difference-in-differences observational study on the spotting capabilities of human users versus the prolific Skycam robot, a Raspberry Pi webcam which automatically submits helicopter sightings from its perch on a downtown window. I figured that, even without users’ names included in the anonymized data table, I could isolate the longitudinal and latitudinal coordinates of the Skycam’s city block to find its spots. But in anonymizing the data, Helicopters of D.C. instead grouped some of the sightings into hyper-specific coordinate pairs, and it would have been near-impossible to isolate Skycam from the human users. 

(An example of the anonymizing job: the Capitol Heights metro station, at the coordinates 38.88973846, -76.91339493, appears 37 times in the data set.)

```{r}
copterspotter_clean |>
  filter(Lat == "38.88973846") |>
  count() |>
  knitr::kable()
```

I pivoted to a before-and-after observational study on the number of helicopter sightings per day, in which the date of each Norton bill’s introduction in Congress served as the treatment, and the number of daily sightings was the dependent variable.


Results:


```{r}
copterspotter_clean_compiled <- copterspotter_clean|>
  count(Date)

copterspotter_clean_compiled |>
  ggplot(aes(x = Date, y = n)) +
  geom_point() +
  geom_smooth() +
    geom_smooth(method = "lm", color = "green", se = FALSE) +
  ylab("Number of Helicopter Spots Per Day") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-03-23"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-05-13"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-07-27"))), color = "indianred")

```
And the data faceted to three common types of helicopter:

The VH3D, the presidential fleet "Sea King":

```{r}
copterspotter_VH3D <- copterspotter_clean |>
  filter(Type == "VH3D")

copterspotter_VH3D_vis <- copterspotter_VH3D |>
  count(Date)

copterspotter_VH3D_vis |>
  ggplot(aes(x = Date, y = n)) +
  geom_point() +
  geom_smooth() +
  ylab("Number of Presidential Copter Spots Per Day") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-03-23"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-05-13"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-07-27"))), color = "indianred") 
```

The Coast Guard's MH65:

```{r}
copterspotter_MH65 <- copterspotter_clean |>
  filter(Type == "MH65")

copterspotter_MH65_vis <- copterspotter_MH65 |>
  count(Date)

copterspotter_MH65_vis |>
  ggplot(aes(x = Date, y = n)) +
  geom_point() +
  geom_smooth() +
  ylab("Number of Coast Guard Copter Spots Per Day") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-03-23"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-05-13"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-07-27"))), color = "indianred") 
```

The UH1N, used by the Air Force First Squadron:

```{r}
copterspotter_UH1N <- copterspotter_clean |>
  filter(Type == "UH1N")

copterspotter_UH1N_vis <- copterspotter_UH1N |>
  count(Date)

copterspotter_UH1N_vis |>
  ggplot(aes(x = Date, y = n)) +
  geom_point() +
  geom_smooth() +
  ylab("Number of Air Force Copter Spots Per Day") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-03-23"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-05-13"))), color = "indianred") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2022-07-27"))), color = "indianred") 
```

The treatment, indicated by red vertical lines, was the introduction of anti-helicopter legislation by Congresswoman Norton. Although none of the bills have passed more than a year later, the hypothesis was that helicopter operators would be sufficiently nervous about the possibility of the bills passing that they would reduce the kind of obnoxious activity that helicopter-watchers would be more likely to observe and report. 

The regression (in both the regression line and smoothed curve metrics) indicates some relationship between the introduction of anti-helicopter legislation and decreased sightings. It's difficult to tell concretely with a small data set -- we can see the regression line driven upwards by forty- and fifty-sighting days. Human error also definitely plays a role. I'm using the quantified observations of helicopter enthusiasts as a stand-in for material that even my congresswoman can't access from the FAA. 

But the faceted data also indicates which helicopter operators are most likely to be affected by the treatment. I chose three common ones: the Marine HMX-1 fleet, which primarily flies VH-3D Sea Kings; the Coast Guard's MH-65 Dolphins; and the Air Force First Helicopter Squadron's UH-1N Hueys. Of the three, the Hueys (which one day in March 2022 were spotted more than twenty times) most closely matched the overall pattern of sighting trends, and showed the most impact from the treatment. 


Conclusion:

Helicopter sightings in 2022 indicate that after each introduction of Congressional legislation to regulate their use, the number of "obnoxious" daily copter sightings fell. For lack of FAA and agency data on the "actual" (as opposed to observer submitted) altitudes and frequencies of helicopter flights, the use of stand-in data paints an imperfect picture of how often obnoxious helicopters flew in D.C. Immediately after the application of the "treatment" effect (the announcement of new legislation), there was an overall downturn in the number of reported helicopter sightings, indicating a possible short-term efficacy of the treatment. Official FAA data or written records would provide better evidence for the efficacy (or lack thereof) of legislation treatment, but while those remain out of public reach, user-submitted data makes for a fairly useful substitute.

